---
title: "Anomaly Detection"
author: "Geoffrey Chege"
date: '2022-06-10'
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction

## Defining the question

- I am a Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).

- I am expected to find out if there are any anomalies in the data.

## Metric for success

- Be able to effectively detect anomalies in the products.

## Understanding the context

- Carrefour operates different store formats, as well as multiple online offerings to meet the growing needs of its diversified customer base.
- In line with the brand’s commitment to provide the widest range of quality products and value for money, Carrefour offers an unrivalled choice of more than 500,000 food and non-food products, and a locally inspired exemplary customer experience to create great moments for everyone every day.

## Recording the experimental design

- Problem Definition
- Anomaly Detection
- Provide insights based on my analysis
- Provide recommendations

## Data Relevance

- Link to the dataset: http://bit.ly/CarreFourSalesDataset

# 2. Loading libraries and dataset

```{r}
# Load tidyverse and anomalize
library(tidyverse)
library(anomalize,warn.conflicts = FALSE)
library(tibbletime)
```

- Data:

```{r}
# read data
forecast <- read.csv("C:/Users/user/Downloads/Supermarket_Sales_Forecasting - Sales.csv")
head(forecast)
```

# 3. Anomaly Detection

## Overview

- We are to check whether there are any anomalies in the given sales dataset. The objective of this task being fraud detection.

```{r}
# checking the structure of our data
str(forecast)
```


```{r}
# checking the shape
dim(forecast)
```

- We have 1000 observations and 2 variables.

```{r}
# converting variables to our preferred format
forecast$Date <- as.Date(forecast$Date, "%m/%d/%Y")
```


```{r}
str(forecast)
```

## Visualization

```{r}
# visualizing our sales
hist(forecast$Sales,col="blue")
```



```{r}
# Sales distribution over time
library(ggplot2)
ggplot(data = forecast, aes(x = Date, y = Sales)) +
      geom_bar(stat = "identity", fill = "green") +
      labs(title = "Sales distribution",
           x = "Date", y = "Sales(ksh)")
```


```{r}
# Ordering the data by Date
forecast = forecast %>% arrange(Date)
head(forecast)
```


```{r}
# Since our data has many records per day,
# We get the average per day, so that the data
forecast = aggregate(Sales ~ Date , forecast , mean)
head(forecast)
```



```{r}

# tbl_time have a time index that contains information about which column

# should be used for time-based subsetting and other time-based manipulation,

forecast= tbl_time(forecast, Date) # Converting data frame to a tibble time (tbl_time)
class(forecast)
```

- I will use the following functions to detect and visualize anomalies:

```{r}
forecast %>%
    time_decompose(Sales) %>%
    anomalize(remainder) %>%
    time_recompose() %>%
    plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)
```

# 4. Conclusion

- There were no anomalies detected in the data.