---
title: "Association Rules"
author: "Geoffrey Chege"
date: '2022-06-10'
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction

## Defining the question

- I am a Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).

- I am expected to find out the associations between products.

## Metric for success

- Be able to effectively identify associations between the different products

## Understanding the context

- Carrefour operates different store formats, as well as multiple online offerings to meet the growing needs of its diversified customer base.
- In line with the brand’s commitment to provide the widest range of quality products and value for money, Carrefour offers an unrivalled choice of more than 500,000 food and non-food products, and a locally inspired exemplary customer experience to create great moments for everyone every day.

## Recording the experimental design

- Problem Definition
- Association Analysis
- Provide insights based on my analysis
- Provide recommendations

## Data Relevance

Link to the dataset: http://bit.ly/SupermarketDatasetII

# 2. Installing packages and loading libraries

```{r eval=FALSE, include=TRUE}
# Installing the necessary packages

install.packages(c("arules", "tidyverse"))
```


```{r}
# Loading the libraries

library(arules)
library(tidyverse)
```


# 3. Loading the dataset

```{r}
# Reading the dataset

assos <- read.transactions("C:/Users/user/Downloads/Supermarket_Sales_Dataset II Par 3.csv", sep = ",", rm.duplicates = TRUE)

assos
```


```{r}
# Verifying the object's class

class(assos)
```


```{r}

# Previewing first 5 transactions

inspect(assos[1:5])
```


```{r}
# Getting a summary of the transactions

summary(assos)
```

# 4. Association Rules.

```{r}
# Plotting the most frequent items both with and without setting the support lower limit

options(repr.plot.width = 15, repr.plot.height = 10)

par(mfrow = c(1, 2))

itemFrequencyPlot(assos, topN = 10,col="lightblue", main = "Frequency plot (default)", cex = 1.5, cex.main= 1.75, cex.lab=1.5, cex.axis=1.2)

itemFrequencyPlot(assos, support = 0.1,col="orange", main = "Frequency plot(supp=0.1)", cex = 1.5, cex.main= 1.75, cex.lab=1.5, cex.axis=1.2)
```


```{r}
# Building a model based on association rules using the apriori function
# supp = 0.001, conf = 0.8

rules <- apriori (assos, parameter = list(supp = 0.001, conf = 0.8))
rules
```


```{r}
# Building a model based on association rules using the apriori function
# supp = 0.002, conf = 0.8

rules1 <- apriori (assos, parameter = list(supp = 0.002, conf = 0.8))
rules1
```


```{r}
# Building a model based on association rules using the apriori function
# supp = 0.001, conf = 0.6

rules2 <- apriori (assos, parameter = list(supp = 0.001, conf = 0.6))
rules2
```

- I will use a model with 74 rules.

```{r}
# Observing rules built in our model i.e. first 10 model rules

inspect(rules[1:10])
```


```{r}
# Inspecting the first 5 rules with the highest lift

inspect(head(rules, n = 5, by = "lift"))
```


```{r}
# Inspecting the first 5 rules with the highest confidence

inspect(head(rules, n = 5, by = "confidence"))
```


```{r}
# Looking at the least popular transactions

itm <- itemFrequency(assos, type = "relative")
head(sort(itm), n = 10)
```


```{r}
# We may want to make a promotion to increase the sale of Tea
# Let us look at what people buy after buying tea

tea = subset(rules, subset = lhs %pin% "tea")

# Then order by confidence
tea = sort(tea, by="confidence", decreasing=TRUE)
inspect(tea[1:5])
```


```{r}
# We may want to make a promotion to increase the sale of ground beef
# Let us look at what people buy after buying ground beef

beef = subset(rules, subset = lhs %pin% "ground beef")
beef

# Then order by confidence
beef = sort(beef, by="confidence", decreasing=TRUE)
inspect(beef[1:5])
```


# 5. Insights

- The insights that can be made from the analysis are as follows:
  - The three most frequently bought items are mineral water, eggs and spaghetti.
  - The 3 least frequently bought items are water spray, napkins and cream. Tea is also among the least frequently purchased items.
  - Ground beef, frozen vegetables and pancakes fell off the most frequently bought items list after support was set to 0.1.


# 6. Recommendations

- In light of the above insights, the following recommendations can be made:
  - To increase the sale of tea, there could be a promotion where tea is sold with milk, eggs or cookies.
  - To increase the sale of ground beef, an offer can be given where ground beef us sold with say, a free bottle of mineral water.