---
title: "Dimensionality Reduction & Feature Selection"
author: "Geoffrey Chege"
date: '2022-06-10'
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction

## 1.1 Defining the question

- I am a Data analyst at Carrefour Kenya and I am currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).

## 1.2 Metric for success

- Be able to reduce the dataset to a low dimensional dataset using the t-SNE algorithm or PCA.

## 1.3 Understanding the context

- Carrefour operates different store formats, as well as multiple online offerings to meet the growing needs of its diversified customer base.
- In line with the brand’s commitment to provide the widest range of quality products and value for money, Carrefour offers an unrivalled choice of more than 500,000 food and non-food products, and a locally inspired exemplary customer experience to create great moments for everyone every day.

## 1.4 Recording the experimental design

- Problem Definition.
- Loading the necessary libraries and the dataset.
- Data Cleaning.
- Exploratory Data Analysis:
  - Univariate Analysis.
  - Bivariate Analysis.
- Part 1: Dimensionality Reduction using t-Distributed Stochastic Neighbor Embedding (t-sne).
- Part 2: Feature Engineering using unsupervised learning.
- Recommendations.

## 1.5 Data Relevance

- Link to the dataset: http://bit.ly/SupermarketDatasetII

# 2. Loading the necessary libraries and the dataset.

```{r}

library(ggplot2)
library(Rtsne)
library(e1071)
library(lattice)
library(corrplot)
library(caret)
library(superml)
library(CatEncoders)
library(FSelector)

library(tidyr)
library(magrittr)
library(warn = -1)
library(RColorBrewer)

library(DataExplorer)
library(Hmisc)
library(pastecs)
library(psych)
library(factoextra)
library(dplyr)
library(ggcorrplot)
library(clustvarsel)
library(mclust)
library("cluster")

```



```{r}

df <- read.csv("C:/Users/user/Downloads/Supermarket_Dataset_1 - Sales Data.csv")
head(df)

```

# 3. Data Cleaning.

## Checking the structure of the data.

```{r}

str(df)

```

- For the analysis, I will need to convert the character columns into factors.

## Data Cleaning:

```{r}

df$Invoice.ID <- as.factor(df$Invoice.ID)
df$Branch <- as.factor(df$Branch)
df$Customer.type <- as.factor(df$Customer.type)
df$Gender <- as.factor(df$Gender)
df$Product.line <- as.factor(df$Product.line)
df$Payment <- as.factor(df$Payment)
df$Date <- as.Date(df$Date, format = "%m/%d/%y")

str(df) #confirming the changes
```

- Next, I will check for duplicates:

```{r}
# checking for duplicates
df[duplicated(df), ]
```

- There are no duplicates in the dataset.

- Checking for mssing values:

```{r}
# checking for missing values
colSums(is.na(df))
```

- There are no missing values in the dataset.

### Outliers

- I will use box plots to check for outliers.

#### Boxplot for "unit.price" column

```{r boxplot unit.price, echo=FALSE}
boxplot(df$Unit.price,
main = "unit.price column Boxplot",  # Titling the boxplot
xlab = "unit.price",                # labelling the axis
col = "blue",                  # setting the colour
border = "black",              # setting the colour of the borders
horizontal = TRUE,
notch = TRUE
)
```

- There are no outliers in the "unit.price" column.

#### Boxplot for "Tax" column

```{r boxplot Tax, echo=FALSE}
boxplot(df$Tax,
main = "Tax column Boxplot",  # Titling the boxplot
xlab = "Tax",                # labelling the axis
col = "brown",                  # setting the colour
border = "black",              # setting the colour of the borders
horizontal = TRUE,
notch = TRUE
)
```

- There are outliers in the "Tax" column. I will keep them in my analysis because they represent true values in the data.

#### Boxplot for "Cogs" column

```{r boxplot cogs, echo=FALSE}
boxplot(df$cogs,
main = "Cogs column Boxplot",        # Titling the boxplot
xlab = "Cogs",                # labelling the axis
col = "orange",              # setting the colour
border = "black",            # setting the colour of the borders
horizontal = TRUE,
notch = TRUE
)
```

- There are outliers in the "Cogs" column. I will keep them in my analysis because they represent true values in the data.

#### Boxplot for "gross.margin.percentage" column

```{r gross.margin.percentage, echo=FALSE}
boxplot(df$gross.margin.percentage,
main = "gross.margin.percentage Boxplot",      # Titling the boxplot
xlab = "gross.margin.percentage",              # labelling the axis
col = "red",                                # setting the colour
border = "black",                           # setting the colour of the borders
horizontal = TRUE,
notch = TRUE
)
```

- There are no outliers in the "Tax" column.

#### Boxplot for "gross.income" column

```{r gross.income, echo=FALSE}
boxplot(df$gross.income,
main = "gross.income Boxplot",      # Titling the boxplot
xlab = "gross.income",              # labelling the axis
col = "green",                        # setting the colour
border = "black",                   # setting the colour of the borders
horizontal = TRUE,
notch = TRUE
)
```

- There are outliers in the "gross.income" column. I will keep them in my analysis because they represent true values in the data.

#### Boxplot for "Rating" column

```{r Rating, echo=FALSE}
boxplot(df$Rating,
main = "Rating Boxplot",      # Titling the boxplot
xlab = "Rating",              # labelling the axis
col = "purple",                                # setting the colour
border = "black",                           # setting the colour of the borders
horizontal = TRUE,
notch = TRUE
)
```

- There are no outliers in the "Rating" column.

#### Boxplot for "Total" column

```{r Total, echo=FALSE}
boxplot(df$Total,
main = "Total Boxplot",      # Titling the boxplot
xlab = "Total",              # labelling the axis
col = "yellow",                                # setting the colour
border = "black",                           # setting the colour of the borders
horizontal = TRUE,
notch = TRUE
)
```

- There are outliers in the "Total" column. I will keep them in my analysis because they represent true values in the data.

# 4. Exploaratory Data Analysis.

## 4.1 Univariate Analysis.

### Distributions

#### Histograms:

```{r}
plot_histogram(df)
```

- From the histograms, we get the following insights:
  - Cogs, gross.income, Tax and Total columns are positively skewed, meaning we expect the mean will be greater than the median.
  - Unit.price and Rating columns have fairly even distribution.

#### Bar Plots:

```{r}
plot_bar(df)
```

- From the bar plots, we observe that Branch, Customer.type, Gender, Product.line and Payment columns have an even distribution.
 
### Description and Summary of Data:

- Description:

```{r}

describe(df)

```

- Summary:

```{r}

summary(df)

```

### A function to get the mode

```{r}
# a function for code
mode <- function(x){
  uniqx <- unique(x)
  uniqx[which.max(tabulate(match(x, uniqx)))]
}
```

#### Unit.price Column

- From the summary and description, we can gather the following about the Unit.price column:
  - Mean: 55.67
  - Median: 55.23
  - Skewness: 0.01
  - Kurtosis: -1.22

- The mode is:

```{r}
mode(df$Unit.price)
```

#### Quantity Column

- From the summary and description, we can gather the following about the Quantity column:
  - Mean: 5.51
  - Median: 5.00
  - Skewness: 0.01
  - Kurtosis: -1.22

- The mode is:

```{r}
mode(df$Quantity)
```

#### Tax Column

- From the summary and description, we can gather the following about the Tax column:
  - Mean: 15.3794
  - Median: 12.0880
  - Skewness: 0.89
  - Kurtosis: -0.09

- The mode is:

```{r}
mode(df$Tax)
```

#### Cogs column

- From the summary and description, we can gather the following about the Cogs column:
  - Mean: 307.59
  - Median: 241.76
  - Skewness: 0.89
  - Kurtosis: -0.09

- The mode is:

```{r}
mode(df$cogs)
```

#### gross.income column

- From the summary and description, we can gather the following about the gross.income column:
  - Mean: 15.3794
  - Median: 12.0880
  - Skewness: 0.89
  - Kurtosis: -0.09

- The mode is:

```{r}
mode(df$gross.income)
```

#### Rating column

- From the summary and description, we can gather the following about the Rating column:
  - Mean: 6.973
  - Median: 7.000
  - Skewness: 0.01
  - Kurtosis: -1.16
 
- The mode is:

```{r}
mode(df$Rating)
```

#### Total column

- From the summary and description, we can gather the following about the Total column:
  - Mean: 322.97
  - Median: 253.85
  - Skewness: 0.89
  - Kurtosis: -0.09
 
- The mode is:

```{r}
mode(df$Total)
```

## 4.2 Bivariate Analysis

```{r product line vs Total Price}
ggplot(df, aes(x=Product.line, y=Total)) +
  geom_point()
```

- Fashion Accessories have the highest Total prices while health and beauty products have lower prices.

```{r Gendered Expenses}
ggplot(df ,aes(Gender, Total)) +
  geom_point()
```

- Total Price is equally distributed in terms of gender

```{r Payment vs Total Price}
ggplot(df, aes(Payment, Total)) +
  geom_point()
```

- The payment methods are nearly identical for the total prices of items at checkouts, with Credit card payments being used for more expensive transactions.

```{r Gross Income vs Total}
ggplot(df, aes(gross.income, Total)) +
  geom_point()
```

- There is a positive linear relationship between the total at checkout and the consumers gross income.

```{r Customer type vs Total}
ggplot(df, aes(Customer.type , Total)) +
  geom_point()
```

- Members and non members have a nearly equal distribution in expenditure with Members having no breaks in prices.

```{r Tax vs Total}
ggplot(df, aes(Tax, Total)) +
  geom_point()
```

- There is a positive linear relationship between tax and total price. As expected, the higher the tax on items, the more they cost.

```{r Unit Price vs Total}
ggplot(df, aes(Unit.price, Total)) +
  geom_point()
```

- There are several positive linear relationships with the Unit Price variable: the higher it is the higher the total price is.

### Heatmap

```{r heatmap}
# Heat map
# Checking the relationship between the variables


# Using Numeric variables only
numeric_tbl <- df %>%
  select_if(is.numeric) %>%
  select(Unit.price, Tax, cogs, gross.income, Rating, Total)

# Calculate the correlations
corr <- cor(numeric_tbl, use = "complete.obs")
ggcorrplot(round(corr, 2),
           type = "full", lab = T)
```

- The following variables show strong correlation:
  - Unit Price with Tax, cogs, gross.income and Total.Strong correlation of 0.63.
  - Tax with cogs, gross.income and Total. Very strong correlation of 1.
  - gross.income to Unit.price, Tax, cogs and Total. Very strong correlation of 1.

# 5. Part 1: Dimensionality Reduction

- This section of the project entails reducing the dataset to a low dimensional dataset using the t-SNE algorithm or PCA. I will perform analysis and provide insights gained.

```{r Label Encoding categorical columns}

# Label Encoding branch column and storing in a copy
branch <- LabelEncoder.fit(df$Branch)
df$Branch <- transform(branch, factor(df$Branch))

# Label Encoding Gender column and storing in a copy
gender <- LabelEncoder.fit(df$Gender)
df$Gender <- transform(gender, factor(df$Gender))

# Label Encoding Customer.type column and storing in a copy
customer <- LabelEncoder.fit(df$Customer.type)
df$Customer.type <- transform(customer, factor(df$Customer.type))

# Label Encoding product.line column and storing in a copy
product <- LabelEncoder.fit(df$Product.line)
df$Product.line <- transform(product, factor(df$Product.line))

# Label Encoding payment column and storing in a copy
pay <- LabelEncoder.fit(df$Payment)
df$Payment <- transform(pay, factor(df$Payment))

```


```{r}

# for plotting

colors = rainbow(length(unique(df$Total)))
names(colors) = unique(df$Total)

```

```{r building the model}

# Executing the algorithm on curated data
model <- Rtsne(df, dims=2, perplexity=30, verbose= TRUE, max_iter=500)

# getting the duration of the execution

exeTimeTsne <- system.time(Rtsne(df, dims = 2, perplexity=30, verbose=TRUE, max_iter = 500))

summary(model)
```

```{r viewing output}
head(model$Y)
```

```{r visualizing results}

plot(model$Y, t='n', main="Output of TSNE")
text(model$Y, labels=df$Total, col=colors[df$Total] )

```



# 6. Part 2: Feature Selection

## 6.1 Filter Methods

```{r}
# Using Numeric variables only
numeric_table <- df %>%
  select_if(is.numeric) %>%
  select(Unit.price, Tax, cogs, gross.income, Rating, Total)
```



```{r correlation method}

corrMat <- cor(numeric_table)

# highly correlated features
high <- findCorrelation(corrMat, cutoff = 0.75)

# names of highly correlated features
names(numeric_table[, high])
```

```{r}
#  Removing Tax, cogs and gross.income
numeric_table2 <- df %>%
  select_if(is.numeric) %>%
  select(Unit.price, Rating, Total)
```



```{r comparison after dropping high}

# data set without highly correlated variables
c2 <- numeric_table[-high]

# plotting
par(mfrow = c(1, 2))

corrplot(corrMat, order = "hclust")
corrplot(cor(c2), order = "hclust")
```

## 6.2 Feature Ranking

```{R}
# From the FSelector package, we use the correlation coefficient as a unit of valuation. 
# This would be one of the several algorithms contained 
# in the FSelector package that can be used rank the variables.
# ---
# 
Scores <- linear.correlation(Total~.,numeric_table)
Scores

```


```{R}
# From the output above, we observe a list containing 
# rows of variables on the left and score on the right. 
# In order to make a decision, we define a cutoff 
# i.e. suppose we want to use the top 5 representative variables, 
# through the use of the cutoff.k function included in the FSelector package. 
# Alternatively, we could define our cutoff visually 
# but in cases where there are few variables than in high dimensional datasets.
# 
# cutoff.k: The algorithms select a subset from a ranked attributes. 
# ---
#
Subset <- cutoff.k(Scores, 4)
as.data.frame(Subset)

```


```{R}
# We could also set cutoff as a percentage which would indicate 
# that we would want to work with the percentage of the best variables.
# ---
#
Subset2 <-cutoff.k.percent(Scores, 0.4)
as.data.frame(Subset2)

```



```{R}
# Instead of using the scores for the correlation coefficient, 
# we can use an entropy - based approach as shown below;
# ---
# 
Scores2 <- information.gain(Total~., numeric_table)

# Choosing Variables by cutoffSubset <- cutoff.k(Scores2, 5)
# ---
# 
Subset3 <- cutoff.k(Scores2, 5)
as.data.frame(Subset3)

```

# 7. Conclusion

- Using Feature Ranking method with information gain of all variables being used as a metric of comparison, the Branch, Customer Type, Gender, Product Line and Unit Price columns would be the best to use for modeling a regressor with respect to Rating.